Total Platform Cyber Protection

### 2016-05-28 paper.tex
Protection against malicious code is added - still working on it

The layers are arraigned from lowest to highest layer. The lower layer
has more control over the computing system, while the upper layers
have less control. For instance, the Hardware layer has total control
over the physical hardware and memory of the machine, while an
application running in the Application layer can only see a portion of
the machine's memory (because of the virtual memory used by the
Operating System).

@@

In our abstract model, the lower levels have more control over the
computing system, yet they have a smaller attacker surface. However,
the lower levels can be attacked by the upper levels. Also, the upper
levels trust the abstractions provided by the lower levels. For
instance, the application-layer visualization of an industrial control
system trusts that the data reported by the hardware layer it
accurate. If an attacker is able to control and manipulate the
hardware to falsely report the state of the system, the entire
integrity of the system can be compromised.

The interconnected nature of the layers means that the security of the
entire layers must be a priority. However, at the same time, the
security improvement of any layer will have an amplifying affect of
the security of the whole system. For instance, improvement to the
security of the application layer decreases the chance of an
application compromise, which further decreases the risk of the attack
propagating throughout the network and through the other layers.

@@

Hardware underpins all of modern computing. While most software is
constructed and written in such as way as to abstract the
peculiarities of the underlying hardware, the software could not
execute without hardware. In our abstracted model of the computing
stack, hardware is at the bottom, which means that it has complete
control of the computing system.

For hardware here, we consider all types of hardware in the computing
system: CPUs, SoCs, or any other integrated circuit (IC) that performs
computation of communication on behalf of the upper computing layers.
This also include networking cards and other peripheral devices
(Bluetooth or USB devices), which may have their own special purpose
embedded hardware.

A key difficult in analyzing the security concerns of hardware is
reverse engineering the workings of the hardware. Traditional
techniques to reverse engineer hardware require destroying the chip,
and if a single hardware component is destroyed to reverse engineer,
there is no guarantee that other hardware is identical to the reverse
engineered chip.

@@

The main threat considered by the literature on hardware is the threat
of Trojan logic. This is malicious logic that is introduced into the
hardware chip by a malicious adversary. Trojan introduction can be
done at several points of the hardware lifecycle:

@@

\item Design. A malicious insider can incorporate a Trojan into the
  design of the hardware.

\item Third-party IP. Oftentimes, hardware designers will take
  advantage of the functionality of third-party components, using them
  as a black-box. Thus, a third-party can include malicious Trojan
  logic.

\item Compilation of Design. When the design of the hardware is
  compiled, but before it is sent for fabrication, the design could be
  changed to incorporate the Trojan logic. Note that this differs from
  Design-time, because the Trojan logic is not included in the design.
  Consider the case of malware on a hardware designer's machine, which
  changes the design when it is compiled.

\item Manufacture. A malicious manufacturer can introduce Trojan logic
  into hardware after the design is received but before the hardware
  is fabricated.

\item Delivery. A malicious hardware component containing Trojan logic
  can be swapped with a benign component after fabrication but before
  delivery and installation.

@@

The malicious Trojan logic behavior comes in three different forms:

@@

\item Trigger-based. The malicious behavior or logic is only expressed
  when certain conditions are met. Consider a Trojan logic in a CPU,
  which changes the executing code to Ring 0 (with the most
  privileges) when a specific sequence of instructions are issued. The
  trigger can be internal to the chip, such as in the CPU case, or
  could be external, for instance a certain radio frequency will
  trigger the malicious code.

\item Constant. The malicious behavior or logic is continually
  executed and subtly changes the output of the hardware in a way that
  compromises security. Consider a Trojan logic of a Wi-Fi card, which
  leaks the Wi-Fi passphrase by embedding it in the operating
  parameters of the Wi-Fi radio.

@@

Of particular practical interest is a description of \emph{actual
  Trojan logic} discovered in a the ProASIC3 Flash FPGA military
chip~\cite{Skorobogatov2012}. While in this case it was not proved
that the logic introduced to the chip was intentionally malicious,
nevertheless there existed functionality in the hardware that was
explicitly denied by the hardware designer.

@@

Another approach is to change the design of the chip to support
runtime monitoring~\cite{Waksman2010}. In this way, the hardware
itself is monitoring itself, rather than external analysis of
side-channels. However, this has all the same drawbacks of changing
the design of hardware, in addition to not being robust in the face of
a powerful attacker who is able to alter the design.

@@

A key problem with hardware security is non-destructive IC
examination, identification, and authentication. How to ensure that
the chip that is designed is what is actually fabricated, with no
additional logic.

@@

Finally, a prevention approach could be taken, to monitoring the upper
layers of the computing stack for attempts to exploit the underlying
hardware. This can also include changing or altering the firmware
layer in such as way that it is not possible to exercise the Trojan
logic, potentially using randomization techniques similar to ASLR.

@@

Firmware is the layer of software that sits and interacts directly
with the hardware. Typically, the firmware is controlled from the
upper layers by device drivers in the Operating System.

Many different hardware contain firmware, for instance the boot BIOS
on a motherboard is a form of firmware, embedded systems often contain
specific firmware, hardware peripheral such as network cards contain
firmware as well. We consider all such areas as firmware.

The key traits of firmware is that, although it is binary code, it is
operating/interacting directly with the hardware, so there are no
OS-style abstraction layers such as virtual memory, well-defined
system calls, or hardware abstraction. Therefore, each firmware is
binary code with is specifically tailed to a specific hardware
platform.

Due to the hardware-specific nature of firmware, \emph{reverse
  engineering} is critical to understanding the firmware. In addition,
reverse engineering is particularly difficult because source code is
frequently unavailable, and, vendors desire to protect their
intellectual property, therefore, they actively obfuscate their
firmware. All this means that, extracting semantically useful
information from firmware is a key first step to performing any
security analysis of the firmware.

@@

Vulnerabilities in the firmware code allow an attacker to execute
arbitrary code with the permissions of the firmware. The class of
vulnerabilities are not unique to firmware, as they are the same class
of vulnerabilities that infect traditional binary software (buffer
overflows, logic flaws, etc.). However, firmware typically has full
control over the entire computing system, therefore vulnerabilities in
firmware are much more severe than those in the application layer. In
addition, if an attacker is able to persist their exploit, because the
compromise occurs at such a low layer of the computing system, the
attacker can persist \emph{even after reinstalling the operating
  system or application}.

The attack surface for firmware varies depending on where it is used.
In the most general case, the attacks can come from the upper layers,
specifically the Operating System or Hypervisor layers, as these are
the layers of the computing stack that can directly communicate to the
firmware. However, other firmware can be attacked either directly
through the Hardware layer itself, or through the data being
transmitted through the firmware from the hardware. For instance, a
network card firmware could be attacked through network traffic that
the firmware processes.

@@

Trojan capabilities in the firmware itself can allow a system to be
remotely controlled or accessed. This is similar to a vulnerability,
in that an attacker is able to control the firmware, however the key
difference lies in the fact that Trojan behavior will be intentionally
obfuscated by an attacker, whereas a vulnerability is accidentally
introduced by a benign developer. Therefore, Trojan behavior is 
more difficult to detect, because there is an active adversary who is
intentionally obfuscating the behavior.

Vulnerabilities in the firmware and Trojan capabilities are related
because both require \emph{understanding} the behavior and semantics
of the firmware code. This is why reverse engineering is such a
critical component of securing the threats against the firmware layer:
in order to detect either vulnerabilities or Trojan functionality, an
automated tool must first understand the semantics of the particular
firmware/hardware combination.

Researchers have attempted to identify real-world attacks on firmware
code~\cite{Tsow2006, Cui2010, Duflot2011, Basnight2013, Cui2013,
  Zaddach2013, Zaddach2013a, Maskiewicz2014, Stuttgen2015}. These
attacks are particularly effective, and researchers have shown that
these are able to persist through a reboot and cannot be detected by
the upper layers.

Other approaches have looked at how to prevent the tampering of
firmware~\cite{Adelstein2002, Zhou2007, Zhou2009}. This protects from
the threat of tampering or altering the firmware, however it does not
address vulnerable firmware.

Another approach has been to include runtime monitoring into the
firmware, so that compromises can be detected~\cite{Cui2010a,
  Cui2011}. This approach is further complicated by the lack of
abstractions in the firmware environment.

Recent work has looked at the problem of updating the design of the
firmware so that it has security features, such as memory handling and
avoiding information leakage~\cite{Koeberl2014, Kauer2007}. However,
forcing firmware developers to use such features is an open issue.

Very few work has looked at how to reverse engineer and understand the
firmware~\cite{Zaddach2014}. This problem is of particular interest,
as understanding the semantics of the firmware and how it interacts
with hardware is a key challenge.

Recent work has looked at how to detect vulnerabilities in
firmware~\cite{Davidson2013, Costin2014, Shoshitaishvili2015}. One of
these has an excellent website\footnote{\url{http://firmware.re}}
that contains the results of their reverse engineering and
vulnerability analysis~\cite{Costin2014}.

Of all the layers, firmware has received the least attention from the
research community. Therefore, we believe that focus on this research
area will yield significant dividends.

The major issue with firmware vulnerabilities is how to decouple
firmware from the hardware, potentially using emulation of the
hardware, in order to aid the analysis of the firmware.

In a similar vein, there is need for automated reverse engineering tools,
including analysis of memory layouts, code modules, even answering the
simple question of what in a firmware binary is code and what is data.

Finally, even if we find vulnerabilities or Trojan logic in a
firmware, there is still the unanswered questions of how to patch or
update the functionality of the firmware while still guaranteeing the
behavior of the firmware. This is particularly complicated due to the
real-time requirements of the firmware and the potentially concurrent
executions.

@@

The Bus and Interconnect is an abstract layer of the computing system
which allows the firmware and hardware to communicate with each other.
Different computing architectures have different technical
implementations of the bus and interconnect. For instance, the
communication network in the automotive system, which has received
significant attention from the research community, uses a CAN bus to
communicate. Modern motherboards us either a Northbound or Southbound
bus to communicate between the CPU and the other peripheral devices,
such as USB.

@@

If the bus and interconnect layer offers advanced functionality, such
as the DMA example, then devices that an attacker is able to connect
to the bus could take advantage of this functionality. In the case of
DMA, this means that any device on the bus has full access to the
physical memory of the system, therefore the attacker has complete
control over the upper layers of the computing system (hypervisor,
operating system, and application).

@@

A major area of research in the Bus and Interconnect layer is work
that shows the danger of a bus and interconnect that allows malicious
devices to either impersonate devices on the bus or leverage bus
capabilities, particularly in the automotive area~\cite{Bonkoski2013,
  Checkoway2011, Hoppe2008, Koscher2010, Rouf2010, Sang2010}.

@@

Another approach is to leverage hardware features of the underlying
bus and interconnect system to reduce the impact of an
attack~\cite{Wolf2012, Wojtczuk2011}. However, as these hardware
functionality must be implemented, it's impact in legacy systems is
limited.

@@

Further research in this area should focus on enabling better
monitoring of the runtime of devices attached to the bus or
interconnect. There is also the need to support logging of the bus or
interconnect activity. This will enable to detection of a malicious
device on the bus or interconnect.

Another promising area of research is into the automated filtering of
the bus communication, potentially allowing an administrator to define
a policy for a specific bus or interconnect system. This would be
particularly effective if it can be implemented on legacy bus and
interconnect systems, with no modification to the devices.

@@

The purpose of the hypervisor layer is to provide virtualization
techniques for the upper layers of the computing stack. In essence,
the Hypervisor layer provides a virtualization or emulation of the
lower layers of the computing stack, so that multiple Operating
Systems, which each assume that the are the sole users of the
underlying hardware, can run concurrently on the same hardware.

For discussing the Hypervisor layer, we use the term \emph{host} to
define the hypervisor itself (which, in some cases can be an Operating
System) and the \emph{guest} is the Operating System that is running
in the hypervisor. A host can run multiple guest Operating Systems.

@@

Other vulnerability defenses include defenses for specific
vulnerability classes, such as buffer overflows~\cite{Dalton2008}.
Randomization techniques attempt to alter the environment of the
operating system in such as way that the system continues to function
but the attacker's exploit fails~\cite{Lin2009, Giuffrida2012}.
Control-flow integrity techniques have been applied to Operating
System code as well~\cite{Li2011a}.

@@

As driver code executes in kernel space, and are a frequent source of
vulnerabilities~\cite{Kadav2012}, researchers have looked into how to
make drivers more robust~\cite{Ganapathy2008, Hruby2012} or isolated
from the rest of the kernel~\cite{Boyd-Wickizer2010}.

@@

We believe that, while significant research has been done on the
Operating System layer, there still remains work to be done on this
area. In particular, as the Operating System is frequently in contact
with malicious applications, it must be hardened from attack.

@@

At the application layer, we consider all the following types of
``applications'' under the application layer umbrella: GUI
applications, command-line applications, server batch process
applications, web applications, or mobile applications. Also, all
programming languages used to develop the applications are also in
scope, from compiled languages such as C or C++ to scripting languages
such as Python and Ruby. In short, all types of computing application
are included in the application layer.

Due to abstractions provided by the lower layers, applications
typically have less privilege than the lower layers of the computing
stack. This is due to the abstractions provided to the application
layer, for instance an application cannot access the underlying hard
drive directly, but must use the Operating System's provided
interface.

Applications developed are typically very complex to understand
completely. This is due in part to the reliance on the abstractions
provided by the lower computing layers (as these abstractions must be
fully understood to understand how the application uses them). Another
aspect is that applications frequently use shared libraries or other
middleware, which as also considered to be part of the application
layer. In this way, multiple different applications can all use the
same underlying library. As the application's behavior depends on the
behavior of \emph{all} used libraries and abstractions, this means
that understanding even a simple application is dependent on the
complexity of the used abstractions.

@@

Applications, by definition must be accessible by users in
order to perform their useful computation. Due to this accessibility,
the application layer, of all the other computing layers, is the entry
point for attacks or exploits.

At a high level the threats to the application layers is an attacker
able to influence the system to perform computation of their choosing.
There are two main vectors for an attacker to influence the system:
(1) software vulnerabilities in an application that can allow an
attacker to exploit the vulnerability to either perform arbitrary
computation or otherwise influence the application and (2) a malicious
application (typically described as a Trojan application) that the
attacker tricks the user into installing and running on the system,
usually by using social engineering techniques.

Vulnerabilities are flaws or bugs in the application (or any software
for that matter) that allows an attacker to compromise the security of
the application. The key difference from normal flaws or bugs in
software is that vulnerabilities allow the program to deviate from the
behavior. These vulnerabilities range the gamut of possible causes
such as programming errors or configuration mistakes. They can be in
the actual, custom application software or in any of the libraries
that the application uses. Some vulnerabilities are specific to a
given programming language, while others are more broad and can any
affect application. Examples of common vulnerabilities are: buffer
overflow, heap overflow, format string, off-by-one, integer overflow,
time-of-check-time-of-use, cross-site scripting, SQL injection, and
logic flaws.

@@

The idea behind this subarea of application layer security research is
to reduce the attack surface of the application. This line of research
views vulnerabilities as a probability, and by reducing the attack
surface, the odds of an attacking being able to exploit a
vulnerability successfully is reduced as well.

@@

Several research works attempt to reduce the attack surface. Some
approaches try to statically analyzing for deadcode, then restrict the
execution of the deadcode during runtime~\cite{Kurmus2011}. Other ways
attempt to leverage compile-time configurability to reduce the attack
surface~\cite{Tartler2012, Kurmus2013, Ruprecht2014}. A different
approach is to reduce the attack surface by partitioning the
application into different privilege levels~\cite{Barth2010, Xu2012}.
Other approaches attempt to restrict the communication between
applications, in order to reduce opportunities for the attacker to
input data to the system~\cite{Kantola2012}.

@@

Modern attackers, exploiting binary applications, such as browsers,
media players, or PDF readers, attempt to hijack the control-flow of
the application. If an attacker successfully hijacks the control-flow
of the application, then she has complete control over the execution
of the application. Many vulnerability classes enable control-flow
hijacking, specifically memory corruption vulnerabilities such as
buffer overflows, heap overflows, and format strings~\cite{Szekeres2013}.

@@

Researchers have advanced the state of attacks against control-flow
hijacking. The most advanced of these is return-oriented programming
(ROP), in which an attacker reuses bits of code from the original
application, so that the attacker can hijack the control-flow of the
application without introducing new code~\cite{Roemer2012,
  Bletsch2010, Tran2011}. Researchers have also improved these ROP
techniques~\cite{Checkoway2010, Buchanan2008, Kayaalp2012, Chen2012,
  Goktas2014, Schwartz2011} and applied them to different
architecture~\cite{Davi2012}.

>>>>

% Sai: should this be 'arranged' instead of 'arraigned'?%Faris: I think it should be 'arranged'.
The layers are arraigned from lowest to highest layer. The lower layer
has more control over the computing system, while the upper layers
have less control. For instance, the Hardware layer has total control
over the physical hardware and memory of the machine, while an
application running in the Application layer can only see a portion of
the machine's memory (because of the virtual memory used by the
Operating System).

@@

In our abstract model, the lower levels have more control over the
computing system, yet they have a smaller attacker surface. However,
the lower levels can be attacked by exploiting the upper levels. Also, the upper
levels trust the abstractions provided by the lower levels. For
instance, the application-layer visualization of an industrial control
system trusts that the data reported by the hardware layer is
accurate. If an attacker is able to control and manipulate the
hardware to falsely report the state of the system, the entire
integrity of the system can be compromised.

The interconnected nature of the layers means that the security of the
entire layer must be a priority. However, at the same time, the
security improvement of any layer will have an amplifying effect on
the security of the whole system. For instance, improvement to the
security of the application layer decreases the chance of an
application compromise, which further decreases the risk of the attack
propagating throughout the network and through the other layers.

@@

Hardware underpins all of modern computing. While most software is
constructed and written in such a way as to abstract the
peculiarities of the underlying hardware, the software cannot
execute without hardware. In our abstracted model of the computing
stack, hardware is at the bottom, which means that it has complete
control of the computing system.

For hardware here, we consider all types of hardware in the computing
system: CPUs, SoCs, or any other integrated circuit (IC) that performs
computation or communication on behalf of the upper computing layers.
This also includes networking cards and other peripheral devices
(Bluetooth or USB devices), which may have their own special purpose
embedded hardware.

A key difficulty in analyzing the security concerns of hardware is
reverse engineering the workings of the hardware. Traditional
techniques to reverse engineer hardware require destroying the chip,
and if a single hardware component is destroyed to reverse engineer,
there is no guarantee that other hardware is identical to the reverse
engineered chip.

@@

The main threat considered by the literature on hardware is the threat
of Trojan logic. This is malicious logic that is introduced into the
% adversary, by definition is a malicious entity.
hardware chip by an adversary. Trojan introduction can be
done at several points of the hardware lifecycle:

@@

\item \textbf{Design}. A malicious insider can incorporate a Trojan into the
  design of the hardware.

\item \textbf{Third-party IP}. Oftentimes, hardware designers will take
  advantage of the functionality of third-party components, using them
  as a black-box. Thus, a third-party can include malicious Trojan
  logic.

\item \textbf{Compilation of Design}. When the design of the hardware is
  compiled, but before it is sent for fabrication, the design could be
  changed to incorporate the Trojan logic. Note that this differs from
  Design-time, because the Trojan logic is not included in the design.
  Consider the case of malware on a hardware designer's machine, which
  changes the design when it is compiled.

\item \textbf{Manufacture}. A malicious manufacturer can introduce Trojan logic into hardware after the design is received but before the hardware
is fabricated.

\item \textbf{Delivery}. A malicious hardware component containing Trojan logic can be swapped with a benign component after fabrication but before
delivery and installation.

@@

The malicious Trojan logic behavior comes in three different forms:
% Sai: is this three or two??

@@

\item \textbf{Trigger-based}. The malicious behavior or logic is only expressed when certain conditions are met. Consider a Trojan logic in a CPU,
  which changes the executing code to Ring 0 (with the most
  privileges) when a specific sequence of instructions are issued. The
  trigger can be internal to the chip, such as in the CPU case, or
  could be external, for instance a certain radio frequency will
  trigger the malicious code.

\item \textbf{Constant}. The malicious behavior or logic is continually
  executed and subtly changes the output of the hardware in a way that
  compromises security. Consider a Trojan logic of a Wi-Fi card, which
  leaks the Wi-Fi passphrase by embedding it in the operating
  parameters of the Wi-Fi radio.

@@

Of particular practical interest is a description of \emph{actual
  Trojan logic} discovered in the ProASIC3 Flash FPGA military
chip~\cite{Skorobogatov2012}. While in this case it was not proved
that the logic introduced to the chip was intentionally malicious,
nevertheless there existed functionality in the hardware that was
explicitly denied by the hardware designer.

@@

Another approach is to change the design of the chip to support
runtime monitoring~\cite{Waksman2010}. In this way, the hardware
itself is monitoring itself, 
% Sai: can we rephrase this^ ? sounds weird %Faris : Removing the first 'itself' may solve the problem.
rather than external analysis of
side-channels. However, this has all the same drawbacks of changing
the design of hardware, in addition to not being robust in the face of
a powerful attacker who is able to alter the design.

@@

A key problem with hardware security is non-destructive IC
examination, identification, and authentication - how to ensure that
the chip that is designed is what is actually fabricated, with no
additional logic.

@@

Finally, a prevention approach could be taken, to monitoring the upper
layers of the computing stack for attempts to exploit the underlying
hardware. This can also include changing or altering the firmware
layer in such a way that it is not possible to exercise the Trojan
logic, potentially using randomization techniques similar to ASLR.

@@

%Note: I did not remove the previous versions of those paragraphs
%(commented them) so feel free to check the previous versions and 
%revert them if my changes have problems.

%Firmware is the layer of software that sits and interacts directly
%with the hardware. Typically, the firmware is controlled from the
%upper layers by device drivers in the Operating System.

%Faris:
Firmware is the layer of software that sits and interacts directly
with the hardware. Typically, the firmware is controlled by the
device drivers in the upper layers of the Operating System.

%Many different hardware contain firmware, for instance the boot BIOS
%on a motherboard is a form of firmware, embedded systems often contain
%specific firmware, hardware peripherals such as network cards contain
%firmware as well. We consider all such areas as firmware.

%Faris:
Many different hardware contain firmware. For instance: the boot BIOS
on a motherboard, embedded systems that contain
specific firmware, hardware peripherals such as network cards. 
We consider all such areas as firmware.

%The key traits of firmware is that, although it is binary code, it is
%operating/interacting directly with the hardware, so there are no
%%OS-style abstraction layers such as virtual memory, well-defined
%system calls, or hardware abstraction. Therefore, each firmware is
%binary code which is specifically tailored to a specific hardware
%platform.

%Faris:
Although firmware is binary code, it is
operating/interacting directly with the hardware, so there are no
OS-style abstraction layers such as virtual memory, well-defined
system calls, or hardware abstraction which is the key trait of firmware. 
Therefore, each firmware is specifically tailored to a specific hardware platform.

%Due to the hardware-specific nature of firmware, \emph{reverse
%	engineering} is critical to understanding the firmware. In addition,
%reverse engineering is particularly challenging because source code is
%frequently unavailable, and vendors' desire to protect their
%intellectual property makes them actively obfuscate their
%firmware. All this means that, extracting semantically useful
%information from firmware is a key first step to performing any
%security analysis of the firmware.

%Faris:
Due to the hardware-specific nature of firmware, \emph{reverse
  engineering} is critical to understand it. In addition,
reverse engineering is particularly challenging because the source code is
frequently unavailable, and vendors' desire to protect their
intellectual property makes them actively obfuscate their
firmware. This means, extracting semantically useful
information from firmware is the first and key step to perform any
security analysis of the firmware.

@@

%Vulnerabilities in the firmware code allow an attacker to execute
%arbitrary code with the permissions of the firmware. The class of
%vulnerabilities are not unique to firmware, as they are the same class
%of vulnerabilities that infect traditional binary software (buffer
%overflows, logic flaws, etc.). However, firmware typically has full
%control over the entire computing system, therefore vulnerabilities in
%firmware are much more severe than those in the application layer. In
%addition, if an attacker is able to persist their exploit --- because the
%compromise occurs at such a low layer of the computing system --- the
%attacker can persist \emph{even after reinstalling the operating
%	system or application}.

%Faris:
Vulnerabilities in the firmware code allow an attacker to execute
arbitrary code with the permissions of the firmware. The class of
vulnerabilities are not unique to firmware, as they are the same class
of vulnerabilities that infect traditional binary software (buffer
overflows, logic flaws, etc.). However, firmware typically has full
control over the entire computing system, therefore vulnerabilities in
firmware are much more severe than those in the application layer. In
addition, the attacker may remain in the system because the
compromise occurs at such a low layer of the computing system so the
exploit persists \emph{even after reinstalling the operating
system or application}.

%The attack surface for firmware varies depending on where it is used.
%In the most general case, the attacks can come from the upper layers,
%specifically the Operating System or Hypervisor layers, as these are
%the layers of the computing stack that can directly communicate with the
%firmware. However, other firmware can be attacked either directly
%through the Hardware layer itself, or through the data being
%transmitted through the firmware from the hardware. For instance, a
%network card firmware could be attacked through network traffic that
%the firmware processes.

%Faris:
The attack surface for firmware varies depending on where it is used.
In the general case, the attacks can come from the upper layers,
specifically the Operating System or Hypervisor layers, as these are
the layers of the computing stack that can directly communicate with the
firmware. However, other firmware can be attacked either directly
through the Hardware layer, or through the data being
transmitted through the firmware from the hardware. For instance, a
network card firmware could be attacked through network traffic that
the firmware processes.

@@

%Trojan capabilities in the firmware itself can allow a system to be
%remotely controlled or accessed. This is similar to a vulnerability,
%in that an attacker is able to control the firmware, however the key
%difference lies in the fact that Trojan behavior will be intentionally
%obfuscated by an attacker, whereas a vulnerability is accidentally
%introduced by a benign developer. Therefore, Trojan behavior is 
%more difficult to detect, because there is an active adversary who is
%intentionally obfuscating the behavior.

%Faris:
Trojan capabilities in the firmware can allow a system to be
remotely controlled or accessed. This is similar to a vulnerability,
in that an attacker is able to control the firmware, however the key
difference lies in the fact that Trojan behavior can be intentionally
obfuscated by an attacker, whereas a vulnerability can accidentally
introduced by a benign developer. Therefore, Trojan behavior is 
difficult to detect.

%Vulnerabilities in the firmware and Trojan capabilities are related
%because both require \emph{understanding} the behavior and semantics
%of the firmware code. This is why reverse engineering is such a
%critical component of securing the threats against the firmware layer:
%in order to detect either vulnerabilities or Trojan functionality, an
%automated tool must first understand the semantics of the particular
%firmware/hardware combination.

%Faris:
Vulnerabilities in the firmware and Trojan capabilities are related
because both require \emph{understanding} the behavior and semantics
of the firmware code. This is why reverse engineering is such a
critical component of securing the firmware layer against the threats.
In order to detect vulnerabilities or trojan functionality, an
automated tool must understand the semantics of the particular
firmware/hardware combination.

\subsection{Research}
%Faris: Some fixes below. 
Researchers have attempted to identify real-world attacks on firmware
code~\cite{Tsow2006, Cui2010, Duflot2011, Basnight2013, Cui2013,
  Zaddach2013, Zaddach2013a, Maskiewicz2014, Stuttgen2015}. These
attacks are particularly effective, and researchers have shown that
these are able to persist through a reboot and cannot be detected by
the upper layers.

Other approaches are about how to prevent the tampering of
firmware~\cite{Adelstein2002, Zhou2007, Zhou2009}. These protect from
the threat of tampering or altering the firmware, however it does not
address vulnerable firmware.

Another approach includes runtime monitoring into the
firmware, so that compromises can be detected~\cite{Cui2010a,
  Cui2011}. This approach is further complicated by the lack of
abstractions in the firmware environment.

Recent work has looked into the problem of updating the design of the
firmware so that it has security features, such as memory handling and
avoiding information leakage~\cite{Koeberl2014, Kauer2007}. However,
forcing firmware developers to use such features is an open issue.

Very few work has looked into how to reverse engineer and understand the
firmware~\cite{Zaddach2014}. This problem is of particular interest,
as understanding the semantics of the firmware and how it interacts
with hardware is a key challenge.

Recent work has looked into how to detect vulnerabilities in
firmware~\cite{Davidson2013, Costin2014, Shoshitaishvili2015}. One of
these has an excellent website\footnote{\url{http://firmware.re}}
that contains the results of their reverse engineering and
vulnerability analysis~\cite{Costin2014}.

\subsection{Suggested Focus Areas}
%Faris: Some fixes below.
Of all the layers, firmware has received the least attention from the
research community. Therefore, we believe that focus on this research
area will yield significant dividends.

The major challenge with preventing firmware vulnerabilities is how to decouple
firmware from the hardware, potentially using emulation of the
hardware, in order to aid the analysis of the firmware.

In a similar vein, there is need for automated reverse engineering tools,
including analysis of memory layouts, code modules, even answering the
simple question of what is code in a firmware binary and what is data.

Finally, even if we find vulnerabilities or Trojan logic in a
firmware, there is still the unanswered question of how to patch or
update the functionality of the firmware while still guaranteeing the
behavior of the firmware. This is particularly complicated due to the
real-time requirements of the firmware and the potentially concurrent
executions.

@@

The Bus and Interconnect is an abstract layer of the computing system
which allows the firmware and hardware to communicate with each other.
Different computing architectures have different technical
implementations of the bus and interconnect. For instance, the
communication network in the automotive system, which has received
significant attention from the research community, uses a CAN bus to
communicate. Modern motherboards use either a Northbound or Southbound
bus --- such as USB --- to communicate between the CPU and other peripheral devices.

@@

If the bus and interconnect layer offers advanced functionality, such
as the DMA example, then the devices that an attacker is able to connect
to the bus could take advantage of this functionality. In the case of
DMA, this means that any device on the bus has full access to the
physical memory of the system, therefore the attacker has complete
control over the upper layers of the computing system (hypervisor,
operating system, and application).

@@

A major area of research in the Bus and Interconnect layer is work
that shows the dangers of a bus and interconnect that allows malicious
devices to either impersonate devices on the bus or leverage bus
capabilities, particularly in the automotive area~\cite{Bonkoski2013,
  Checkoway2011, Hoppe2008, Koscher2010, Rouf2010, Sang2010}.

@@

Another approach is to leverage hardware features of the underlying
bus and interconnect system to reduce the impact of an
attack~\cite{Wolf2012, Wojtczuk2011}. However, as these hardware
functionalities must be implemented, it's impact in legacy systems is
limited.

@@

Further research in this area should focus on enabling better
monitoring of the runtime of devices attached to the bus or
interconnect. There is also the need to support logging of the bus or
interconnect activity. This will enable detection of a malicious
device on the bus or interconnect.

Another promising area of research is the automated filtering of
the bus communication, potentially allowing an administrator to define
a policy for a specific bus or interconnect system. This would be
particularly effective if it can be implemented on legacy bus and
interconnect systems, with no modification to the devices.

@@

The purpose of the hypervisor layer is to provide virtualization
techniques for the upper layers of the computing stack. In essence,
the Hypervisor layer provides a virtualization or emulation of the
lower layers of the computing stack, so that multiple Operating
Systems, which each assume that they are the sole users of the
underlying hardware, can run concurrently on the same hardware.

For discussing the Hypervisor layer, we use the term \emph{host} to
define the hypervisor itself (which, in some cases can be an Operating
System) and the \emph{guest} is the Operating System that is running
on the hypervisor. A host can run multiple guest Operating Systems.

@@

Other vulnerability defenses include defenses for specific
vulnerability classes, such as buffer overflows~\cite{Dalton2008}.
Randomization techniques attempt to alter the environment of the
operating system in such a way that the system continues to function
but the attacker's exploit fails~\cite{Lin2009, Giuffrida2012}.
Control-flow integrity techniques have been applied to Operating
System code as well~\cite{Li2011a}.

@@

% either driver code execute or change `and' to `is'
As driver code executes in kernel space, and is a frequent source of
vulnerabilities~\cite{Kadav2012}, researchers have looked into how to
make drivers more robust~\cite{Ganapathy2008, Hruby2012} or isolated
from the rest of the kernel~\cite{Boyd-Wickizer2010}.

@@

We believe that, while significant research has been done on the
Operating System layer, there still remains work to be done in this
area. In particular, as the Operating System is frequently in contact
with malicious applications, it must be hardened from attack.

@@

At the application layer, we consider all the following types of
``applications'' under the application layer umbrella: GUI
applications, command-line applications, server batch process
applications, web applications, or mobile applications. Also, all
programming languages used to develop the applications are also in
scope, from compiled languages such as C or C++ to scripting languages
such as Python and Ruby. In short, all types of computing applications
are included in the application layer.

Due to abstractions provided by the lower layers, applications
typically have lesser privilege than the lower layers of the computing
stack. This is due to the abstractions provided to the application
layer, for instance an application cannot access the underlying hard
drive directly, but must use the Operating System's provided
interface.

Applications developed are typically very complex to understand
completely. This is due in part to the reliance on the abstractions
provided by the lower computing layers (as these abstractions must be
fully understood to understand how the application uses them). Another
aspect is that applications frequently use shared libraries or other
middleware, which are also considered to be part of the application
layer. In this way, multiple different applications can all use the
same underlying library. As the application's behavior depends on the
behavior of \emph{all} used libraries and abstractions, this means
that understanding even a simple application is dependent on the
complexity of the used abstractions.

@@

Applications, by definition, must be accessible by users in
order to perform their useful computation. Due to this accessibility,
the application layer, of all the other computing layers, is the entry
point for attacks or exploits.

At a high level the threat to the application layer is attackers being
able to influence the system to perform computation of their choosing.
There are two main vectors for an attacker to influence the system:
(1) software vulnerabilities in an application that can allow an
attacker to exploit the vulnerability to either perform arbitrary
computation or otherwise influence the application and (2) a malicious
application (typically described as a Trojan application) that the
attacker tricks the user into installing and running on the system,
usually by using social engineering techniques.

Vulnerabilities are flaws or bugs in the application (or any software
for that matter) that allows an attacker to compromise the security of
the application. The key difference from normal flaws or bugs in
software is that vulnerabilities allow the program to deviate from the expected behavior. These vulnerabilities range the gamut of possible causes
such as programming errors or configuration mistakes. They can be in
the actual, custom application software or in any of the libraries
that the application uses. Some vulnerabilities are specific to a
given programming language, while others are more broad and can any
affect application. Examples of common vulnerabilities are: buffer
overflow, heap overflow, format string, off-by-one, integer overflow,
time-of-check-time-of-use, cross-site scripting, SQL injection, and
logic flaws.

@@

The idea behind this subarea of application layer security research is
to reduce the attack surface of the application. This line of research
views vulnerabilities as a probability, and by reducing the attack
surface, the odds of an attack being able to exploit a
vulnerability successfully is reduced as well.

@@

Several research works attempt to reduce the attack surface. Some
approaches try to statically analyze for deadcode, then restrict the
execution of the deadcode during runtime~\cite{Kurmus2011}. Other ways
attempt to leverage compile-time configurability to reduce the attack
surface~\cite{Tartler2012, Kurmus2013, Ruprecht2014}. A different
approach is to reduce the attack surface by partitioning the
application into different privilege levels~\cite{Barth2010, Xu2012}.
Other approaches attempt to restrict the communication between
applications, in order to reduce opportunities for the attacker to
input data to the system~\cite{Kantola2012}.

@@

Modern attackers, exploiting binary applications such as browsers,
media players, or PDF readers, attempt to hijack the control-flow of
the application. If an attacker successfully hijacks the control-flow
of the application, then she has complete control over the execution
of the application. Many vulnerability classes enable control-flow
hijacking, specifically memory corruption vulnerabilities such as
buffer overflows, heap overflows, and format strings~\cite{Szekeres2013}.

@@

Researchers have advanced the state of attacks against control-flow
hijacking. The most advanced of these is return-oriented programming
(ROP), in which an attacker reuses bits of code from the original
application, so that the attacker can hijack the control-flow of the
application without introducing new code~\cite{Roemer2012,
  Bletsch2010, Tran2011}. Researchers have also improved these ROP
techniques~\cite{Checkoway2010, Buchanan2008, Kayaalp2012, Chen2012,
  Goktas2014, Schwartz2011} and applied them to a different
architecture~\cite{Davi2012}.
